---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# StreamData

<!-- badges: start -->
<!-- badges: end -->

The goal of StreamData is to provide easy access and automated data management to USGS stream BioData, pesticide county use estimates, and measured water quality data.

## Installation
Note: because this package is held within a private repository on GitHub, you will need to use the `auth_token` argument within `install_github()`. Generate a personal access token (PAT) in <https://github.com/settings/tokens> and supply the provided code to this argument.


You can install the released version of `StreamData` from [GitHub](https://github.com/StreamData/StreamData) with:


```
library(devtools)

devtools::install_github("StreamData/StreamData",
                         auth_token = "tokenstring",
                         build_vignette = TRUE
                         )
```

## Example

This is a basic example which shows you how to use one of the *getBioData* set of functions to access the fish dataset.

```{r example}
library(StreamData)

Fish <- getFishData(taxonLevel = "Family")

head(Fish)[,1:35]
```
