---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# finsyncR -this is where the hexsticker will go

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-stable-green.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![R-CMD-check](https://github.com/StreamData/finsyncR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/StreamData/finsyncR/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of finsyncR is to provide easy access and automated data management to USGS and EPA stream BioData, pesticide county use estimates, and measured water quality data.


## Installation
Note: because this package is held within a private repository on GitHub, you will need to use the `auth_token` argument within `install_github()`. Generate a personal access token (PAT) in <https://github.com/settings/tokens> and supply the provided code to this argument.


You can install the released version of `finsyncR` from [GitHub](https://github.com/StreamData/finsyncR) with:


```
library(devtools)

devtools::install_github("StreamData/finsyncR",
                         auth_token = "tokenstring",
                         build_vignette = TRUE)
```

## Example

This is a basic example which shows you how to use one of the *getBioData* set of functions to access the fish dataset.


```{r example}
library(finsyncR)

Fish <- getFishData(taxonLevel = "Species")

head(Fish)[,c(1:5,36,90,400)]
```

## Metadata
Metadata can be found in the "Metadata" vignette: `vignette("Metadata", package = "finsyncR")`.

## Using finsyncR
A compendium on the correct use of the finsyncR package and its output datasets can be found in the "Getting Started" vignette: `vignette("GettingStarted", package = "finsyncR")`.


## Open-Source Code Policy

Effective August 8, 2016, the [OMB Mandate: M-16-21; Federal Source Code Policy: Achieving Efficiency, Transparency, and Innovation through Reusable and Open Source Software](https://obamawhitehouse.archives.gov/sites/default/files/omb/memoranda/2016/m_16_21.pdf) applies to new custom-developed code created or procured by EPA consistent with the scope and applicability requirements of Office of Management and Budget's (OMB's) Federal Source Code Policy. In general, it states that all new custom-developed code by Federal Agencies should be made available and reusable as open-source code.

The EPA specific implementation of OMB Mandate M-16-21 is addressed in the [System Life Cycle Management Procedure](https://www.epa.gov/irmpoli8/policy-procedures-and-guidance-system-life-cycle-management-slcm). EPA has chosen to use GitHub as its version control system as well as its inventory of open-source code projects. EPA uses GitHub to inventory its custom-developed, open-source code and generate the necessary metadata file that is then posted to code.gov for broad reuse in compliance with OMB Mandate M-16-21.

If you have any questions or want to read more, check out the [EPA Open Source Project Repo](https://github.com/USEPA/open-source-projects) and [EPA's Interim Open Source Code Guidance](https://www.epa.gov/developers/open-source-software-and-epa-code-repository-requirements).

## License

All contributions to this project will be released under the CCO-1.0 license file dedication. By submitting a pull request or issue, you are agreeing to comply with this waiver of copyright interest.

## Disclaimer

The United States Environmental Protection Agency (EPA) GitHub project code is provided on an "as is" basis and the user assumes responsibility for its use. EPA has relinquished control of the information and no longer has responsibility to protect the integrity , confidentiality, or availability of the information. Any reference to specific commercial products, processes, or services by service mark, trademark, manufacturer, or otherwise, does not constitute or imply their endorsement, recommendation or favoring by EPA. The EPA seal and logo shall not be used in any manner to imply endorsement of any commercial product or activity by EPA or the United States Government.
